---
title: "30N_rmarkdown - Analyzing trends in Canadian glacier mass"
author: "Me"
date: "9 July 2017"
output:
  html_document:
    toc: yes
    toc_float: yes
    css: 30N_styles.css
  pdf_document:
    keep_tex: yes
    latex_engine: xelatex
    toc: yes
params:
  filename: glacier.csv
bibliography: bibl.bib
urlcolor: blue
---

```{r setup, include=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggvis)
options(scipen = 0, digits = 3)
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE)
```

## Introduction

The data used in study is taken from <http://open.canada.ca/en/open-data>[^1]. The algorithm is taken from @AntonovErmakov_RandomCubaturesQMC.

[^1]: Here is the [direct link](https://raw.githubusercontent.com/tonytonov/Rcourse/master/R%20markdown/demos/glacier.csv) to data download.

```{r}
glacier <- read.csv(paste0("https://raw.githubusercontent.com/tonytonov/Rcourse/master/R%20markdown/demos/", params$filename), na.strings = "..", comment.char = "#")
```

The data set contains `r nrow(glacier)` measurements of `r nlevels(glacier$GEO)` Canadian glacier mass balance, collected from `r min(glacier$Ref_Date)` till `r max(glacier$Ref_Date)`. Namely, the file includes these glaciers:

```{r, echo=TRUE, tidy=TRUE}
levels(glacier$GEO)
```

## Hypothesis

We are interested in finding out whether there is a statistically significant change in mass balance over the observed time period. For these purposes we use **R** (version `r getRversion()`) and an appropriate statistical test called _t-test_: $$t = \frac{\overline{x} - \mu_0}{s/\sqrt{n}} $$

The workflow is as follows:

1. Read file
2. Run t-test for each glacier and collect p-values
3. Support the evidence with 
    + a table of results;
    + a plot that could help demonstrate the effect.

## Analysis

```{r analysis, cache=TRUE}
glacier1 <- glacier %>% 
  select(Ref_Date, GEO, MEASURE, Value) %>% 
  filter(MEASURE == "Annual mass balance") %>% 
  separate(GEO, c("Name", "Location"), sep = " - ")

# descriptive analysis
g1 <- glacier1 %>% 
  group_by(Name) %>% 
  summarise(YearsObserved = n(),
            MeanChange = mean(Value, na.rm = T),
            WorstChange = min(Value, na.rm = T),
            WorstYear = Ref_Date[which.min(Value)])

# t-test
g2 <- glacier1 %>% 
  group_by(Name) %>% 
  do({
    tt <- t.test(.$Value, alternative = "less", mu = 0, conf.level = 0.99)
    data.frame(PValue = tt$p.value,
               ConfidenceLimit = tt$conf.int[2])
  })

# Сливаем две таблицы в одну
left_join(g1, g2, by = "Name") %>% 
  knitr::kable(caption = "Descriptive statistics and confidence intervals",
               digits = c(0, 0, 2, 0, 0, 10, 2))
```

```{r ggplot, fig.width=10, fig.height=7, fig.cap="Annual dynamics of glacier's mass"}
#ggplot(glacier1, aes(Ref_Date, Value))+
#  geom_line()+
#  geom_hline(data = g1, aes(yintercept = MeanChange),
#             color = "red", linetype = "dashed", alpha = 0.8)+
#  facet_wrap(~Name, nrow = 2)
ggvis(glacier1, ~Ref_Date, ~Value, stroke = ~Name) %>% 
  layer_lines()
```

The conducted analysis supports the evidence that Canadian glaciers are <u>melting</u>.

## Bibliography
